Follow-up Samples

### 2017-09-16 submission/manuscript_R2.tex

\textbf{Results.} There were small changes to the bacterial community
associated with adenoma or advanced adenoma and large changes associated
with carcinoma. The communities from patients with carcinomas changed
significantly more than those with adenoma following treatment (P-value
\textless{} 0.001). Although treatment was associated with intrapersonal
changes, the change in the abundance of individual OTUs in response to
treatment was not consistent within diagnosis groups (P-value
\textgreater{} 0.05). Because the distribution of OTUs across patients
and diagnosis groups was irregular, we used the Random Forest machine
learning algorithm to identify groups of OTUs that could be used to
classify pre and post-treatment samples for each of the diagnosis
groups. Although the adenoma and carcinoma models could reliably
differentiate between the pre and post-treatment samples (P-value
treatment models had AUCs that were significantly higher than a random
model permutation (P-value \textless{} 0.001), the advanced-adenoma
model could not (P-value = 0.61). Furthermore, there was little overlap
between the OTUs that were indicative of each treatment. To determine
whether individuals who underwent treatment were more likely to have
OTUs associated with normal colons we used a larger cohort that
contained individuals with normal colons and those with adenomas,
advanced adenomas, and carcinomas. We again built Random Forest models
and measured the change in the positive probability of having one of the
three diagnoses to assess whether the post-treatment samples received
the same classification as the pre-treatment samples. Samples from
patients who had carcinomas changed towards a microbial milieu that
resembles the normal colon after treatment (P-value \textless{} 0.001).
Finally, we were unable to detect any significant differences in the
microbiota of individuals treated with surgery alone and those treated
with chemotherapy or chemotherapy and radiation (P-value \textgreater{}
0.05).

>>>>

\textbf{Results.} There were small changes to the bacterial community
associated with adenoma or advanced adenoma and large changes associated
with carcinoma. The communities from patients with carcinomas changed
significantly more than those with adenoma following treatment (P-value
\textless{} 0.001). Although treatment was associated with intrapersonal
changes, the change in the abundance of individual OTUs in response to
treatment was not consistent within diagnosis groups (P-value
\textgreater{} 0.05). Because the distribution of OTUs across patients
and diagnosis groups was irregular, we used the Random Forest machine
learning algorithm to identify groups of OTUs that could be used to
classify pre and post-treatment samples for each of the diagnosis
groups. Although the adenoma and carcinoma models could reliably
differentiate between the pre and post-treatment samples (P-value
\textless{} 0.001), the advanced-adenoma model could not (P-value =
0.61). Furthermore, there was little overlap between the OTUs that were
indicative of each treatment. To determine whether individuals who
underwent treatment were more likely to have OTUs associated with normal
colons we used a larger cohort that contained individuals with normal
colons and those with adenomas, advanced adenomas, and carcinomas. We
again built Random Forest models and measured the change in the positive
probability of having one of the three diagnoses to assess whether the
post-treatment samples received the same classification as the
pre-treatment samples. Samples from patients who had carcinomas changed
towards a microbial milieu that resembles the normal colon after
treatment (P-value \textless{} 0.001). Finally, we were unable to detect
any significant differences in the microbiota of individuals treated
with surgery alone and those treated with chemotherapy or chemotherapy
and radiation (P-value \textgreater{} 0.05).
